---
title: 'Matrix sampling for TFIDF'
author: "Santina Lin"
date: "November 9, 2016"
output: github_document
---

## Overview 
Looking at how sampling rate affect the variances of the matrix. 

The idea is that, the same number of singular values produce the same variance in a sampled matrix as the one in the original non-sampled matrix. 

In this experiment, we first convert the original term-frequency into TFIDF (term-frequency-inverse-document-frequency) and then multiply them by a factor (10, 100, 1000), sample at various rates (0.01, 0.05, 0.1, 0.2, 0.4) with three replicates [1-3]. Since there are 10 different original matrices, that's a total of 450 sampled matrices. 

We then do SVD on each of the matrices (nsv = 800) and see how the coverage (ratio of total variance for a given NSV to the forbenius norm of the original matrix)

## Data
Setting up the required packages and import the data 
```{r libraries, message=FALSE}

library(ggplot2)
library(plyr) # note to self: load this before dplyr always. 
library(dplyr)
library(magrittr)
```
Get the data
```{r}
tfidf <- read.table("TFIDF.result", header=TRUE)
tfidf_original <- read.table("tfidf_original.summary", header=TRUE)
```
Inspect data 
```{r}
str(tfidf)
str(tfidf_original)
```

## Fixing my original calculation 
On Dec 20, I found that I forgot to take the square root of my frobenius norm and variance (sum of squares). Therefore, I'm doing it now to fix it with in the dataset itself to avoid having to rerun the command. 

```{r}
tfidf <- transform(tfidf, forbenius=sqrt(forbenius), variance=sqrt(variance))
tfidf_original <- transform(tfidf_original, forbenius=sqrt(forbenius), variance=sqrt(variance))
```

Cleaning up the data to extract out information in the names
```{r, message=FALSE}
# Let's separate out the replication and rate  in different columns 
tfidf <- tfidf %>% tidyr::separate(matrix_name, c("matrix_name", "scalar", "rate", "replicate"), sep = "_") %>% readr::type_convert()
# Fix column name of tfidf and do the same thing 
tfidf$matrix_name <- as.factor(gsub(".tfidf", "", tfidf$matrix_name))
tfidf$nsv <- as.factor(tfidf$nsv)

# Fix tfidf_original to prepare for a rbind
tfidf_original$replicate = 1
tfidf_original$rate = 1
tfidf_original$scalar = 1
tfidf_original$matrix_name <- as.factor(gsub(".tfidf", "", tfidf_original$matrix_name))
tfidf_original$nsv <- as.factor(tfidf_original$nsv)


# rbind and set factors as necessary
all <- rbind(tfidf, tfidf_original)
all$scalar <- as.factor(all$scalar)
```

## Calculate coverage 

```{r}
get_summary <- function(data) {
  summary <- data %>% dplyr::group_by(matrix_name, rate, nsv, scalar) %>% dplyr::summarise_at(c("variance", "forbenius"), c("mean", "sd")) %>% dplyr::mutate(cov = variance_mean / forbenius_mean)
}
tfidf_summary <- get_summary(all)
```


## Result 
```{r tfidf_coverage}
# Plot tfidf 
ggplot(tfidf_summary, aes(x = rate, y = cov, colour = nsv, group = nsv)) + geom_point() + facet_wrap(~matrix_name) + theme_bw()
```

## See the difference in scalar

Scalar = 10
```{r tfidf_coverage_scalar10}
# Create different data.frame for different scalar group (including the sample = 1 group)
all_10 <- rbind(tfidf_original, tfidf[tfidf$scalar==10,])
all_10 <- get_summary(all_10)
ggplot(all_10, aes(x = rate, y = cov, colour = nsv, group = nsv)) + geom_point() + geom_line() +  facet_wrap(~matrix_name, nrow=2) + theme_bw()
``` 

Scalar = 100
```{r tfidf_coverage_scalar100}
all_100 <- rbind(tfidf_original, tfidf[tfidf$scalar==100,])
all_100 <- get_summary(all_100)
ggplot(all_100, aes(x = rate, y = cov, colour = nsv, group = nsv)) + geom_point() + geom_line() +  facet_wrap(~matrix_name, nrow=2) + theme_bw() 
```

Scalar = 1000
```{r tfidf_coverage_scalar1000}
all_1000 <- rbind(tfidf_original, tfidf[tfidf$scalar==1000,])
all_1000 <- get_summary(all_1000)
ggplot(all_1000, aes(x = rate, y = cov, colour = nsv, group = nsv)) + geom_point() + geom_line() +  facet_wrap(~matrix_name, nrow=2) + theme_bw() 
```

## Look at error estimates and singular values
```{r erorr_estimate, warning=FALSE}
# Data process. Loading from my own project space because I don't want to commit the result file. 
svals_tfidf <- read.table("/projects/slin_prj/PubMed_Experiment/graph_sparsification/svals_sampled_tfidf.result", header=TRUE)
svals_tfidf <- svals_tfidf %>% dplyr::group_by(matrix_name, rate, multiplier, rank) %>% dplyr::summarise_at(c("singular_value", "error_estimate"), c("mean", "sd"))
svals_tfidf <- svals_tfidf %>% tidyr::separate(matrix_name, c("id"), sep = "[.]") # simplify matrix name 
svals_tfidf$rate <- factor(svals_tfidf$rate)
```

For multiplier = 10 
```{r multiplier_10}

# Excluding 93 to see the others better. 
ggplot(svals_tfidf[c(svals_tfidf$id != "93", svals_tfidf$multiplier==10), ], aes(x=rank, y=error_estimate_mean, colour=rate)) + geom_point(alpha=0.5, size=0.5) + facet_wrap(~id) + scale_color_brewer(palette="Spectral") + ylab("singular values") + theme_bw()
```
From the look of it, the error estimate are similar for matrices sampled at different rate. Powergraph's error of singular value estimates are similar for the same matrix sampled at different rate. 

A look at the singular values 

```{r mean_svals, warning=FALSE}
ggplot(svals_tfidf, aes(x=rank, y=singular_value_mean, colour=rate)) + geom_point(alpha=0.5, size=0.5) + facet_wrap(~id, nrow=2) + scale_color_brewer(palette="Spectral") + ylab("singular value") + theme_bw()

# Exclude 93 for 3 by 3 
ggplot(svals_tfidf[svals_tfidf$id != "93", ], aes(x=rank, y=singular_value_mean, colour=rate)) + geom_point(alpha=0.5, size=0.5) + facet_wrap(~id) + scale_color_brewer(palette="Spectral") + ylab("singular value") + theme_bw()
``` 

## Running time 

Load data
```{r}
run_time <- read.table("tfidf_runningtime.result", header=TRUE)
str(run_time)
run_time$scalar <- factor(run_time$scalar)
run_time$matrix_name <- factor(run_time$matrix_name)
```

Summarize the data by taking the average

```{r}
get_summary <- function(data) {
  summary <- data %>% dplyr::group_by(matrix_name, sample_rate, scalar) %>% dplyr::summarise_at(c("run_time"), c("mean", "sd")) %>% dplyr::mutate(de = sd/sqrt(3) )
}

run_time_summary <- get_summary(run_time)
head(run_time_summary)
```

Graph result 

```{r runningtime_tfidf_summary}
ggplot(run_time_summary, aes(x = sample_rate, y = mean, colour = scalar, group = scalar)) + geom_point() + geom_line() + geom_errorbar(aes(ymax = mean + de, ymin=mean - de), width=0.01) +  facet_wrap(~matrix_name, nrow=2) + theme_bw()
```

Remove 90, 91, and 92 to show the data better. 
```{r runningtime_tfidf_no_outliers}
run_time_summary_subset <- subset(run_time_summary, !matrix_name %in% c("90", "91", "92"))
ggplot(run_time_summary_subset, aes(x = sample_rate, y = mean, colour = scalar, group = scalar)) + geom_point() + geom_line() + geom_errorbar(aes(ymax = mean + de, ymin=mean - de), width=0.01) +  facet_wrap(~matrix_name) + theme_bw()
```


## Sparsity 

Load data
```{r}
sparsity <- read.table("sparsity_TFIDF.result", header=TRUE)
str(sparsity)
sparsity$scalar <- factor(sparsity$scalar)
sparsity$matrix_name <- factor(sparsity$matrix_name)
```

Summarize the data by taking the average

```{r}
get_summary <- function(data) {
  summary <- data %>% dplyr::group_by(matrix_name, sample_rate, scalar) %>% dplyr::summarise_at(c("n_entries"), c("mean", "sd")) %>% dplyr::mutate(de = sd/sqrt(3) )
}

sparsity_summary <- get_summary(sparsity)
head(sparsity_summary)
```

Graph result 

```{r sparsity_tfidf_summary}
ggplot(sparsity_summary, aes(x = sample_rate, y = mean, colour = scalar, group = scalar)) + geom_point() + geom_line() + geom_errorbar(aes(ymax = mean + de, ymin=mean - de), width=0.01) +  facet_wrap(~matrix_name, nrow=2) + theme_bw()
```

Make 3 by 3 
```{r sparsity_tfidf_3by3}
ggplot(sparsity_summary[sparsity_summary$matrix_name!="9", ], aes(x = sample_rate, y = mean, colour = scalar, group = scalar)) + geom_point() + geom_line() + geom_errorbar(aes(ymax = mean + de, ymin=mean - de), width=0.01) +  facet_wrap(~matrix_name) + theme_bw()
```


