---
title: "Analysis of the singular values"
author: "Santina Lin"
date: "December 17, 2016"
output: 
  html_document: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
```

## Data 

The tf matrix was sampled at 0.1 rate and was SVD with nsv=2000. 

We'll look at the log file of the GraphLab output from running collaborative filtering on this matrix. 
```{r}
singular_value <- read.table("sval_tf_r0.1_nsv2000.summary", header=TRUE)
head(singular_value)
``` 

### Error estimates 

First let's plot the error 
```{r error_estimates}
ggplot(singular_value, aes(x = rank, y = error_estimate)) + geom_point(size=1, alpha=0.6) + ylab("Error estimate") + xlab("Rank") 
```

```{r available_nsv}
nsv <- sum(singular_value$error_estimate < 0.5)
nsv
singular_value$error_estimate[nsv]
singular_value$error_estimate[nsv+1]
```

### Singular values 

Plotting the singular value
```{r singular_value_plot}
ggplot(singular_value, aes(x = rank, y = singular_value)) + geom_point(size=1, alpha=0.6) + ylab("Singular value") + xlab("Rank") 
```

As expected in this sanity check, it's a downward curve that drops in tangential slope sharply after the first few singular value. 

### Conclusion

From the error estimation, we'll be using about `r nsv` nsv to run GraphLab on our original big matrix.  

