---
title: "Analysis of the singular values"
author: "Santina Lin"
date: "December 17, 2016"
output: 
  html_document: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
```

## Data 

The tf matrix was sampled at 0.1 rate and was SVD with nsv=2000. 

We'll look at the log file of the GraphLab output from running collaborative filtering on this matrix. 
```{r}
svals_sampled <- read.table("svals_tf_sampled_0.1_nsv2000.txt", header=TRUE)
svals_full <- read.table("svals_tf_nsv2000.txt", header=TRUE)
head(svals_sampled)
``` 

Combine the two data set together 
```{r}
svals_full$matrix <- factor("full matrix")
svals_sampled$matrix <- factor("sampled matrix")
all <- rbind(svals_full, svals_sampled)

```

### Error estimates 

First let's plot the error 
```{r error_estimates}
ggplot(all, aes(x = rank, y = error_estimate, color=matrix)) + geom_point(size=1, alpha=0.6) + ylab("Error estimate") + xlab("Rank") + theme(legend.title=element_blank())
```

```{r available_nsv}
nsv <- sum(svals_sampled$error_estimate < 0.5)
nsv
svals_sampled$error_estimate[nsv]
svals_sampled$error_estimate[nsv+1]
```

### Singular values 

Plotting the singular value
```{r singular_value_plot}
ggplot(all, aes(x = rank, y = singular_value, color=matrix)) + geom_point(size=1, alpha=0.6) + ylab("Singular value") + xlab("Rank") + theme(legend.title=element_blank())
```

As expected in this sanity check, it's a downward curve that drops in tangential slope sharply after the first few singular value. 

### Coverage 

```{r coverage}
coverage_sampled <- read.table("tf_nsv_2000_0.1_coverage.txt", header=TRUE)
coverage_full <- read.table("tf_nsv_2000_coverage.txt", header=TRUE)
coverage_sampled$matrix <- factor("sampled matrix")
coverage_full$matrix <- factor("full matrix")
all <- rbind(coverage_full, coverage_sampled)
ggplot(all, aes(x = nsv, y = coverage, color=matrix)) + geom_point(size=1, alpha=0.6) + ylab("Singular value") + xlab("nsv") + theme(legend.title=element_blank())
```


### Conclusion

From the error estimation, we'll be using about `r nsv` nsv to run GraphLab on our original big matrix.  

And the result is shown as the overlap graph. 

